COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-07-03 main.tex
Fix typos, update appendix label and figure caption

\noindent where $\lambda = \varpi + M$ is the mean longitude of a
body, with $M$ being the mean anomoly. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citep{2000ssd..book.....M}. An important 
point to note, which we will come back to later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of 
the $\phi$. For second and higher order resonances, the maximum libration width is given by equation 8.58 in \citet{2000ssd..book.....M} as

@@

This also suggests that the prominent bumps and dips seen in figure \ref{fig:coll_hist_r} are mainly produced by bodies outside of resonance, although 
the resonant population does make some of the features more pronounced. To further understand this, we consider the nonaxisymmetric structure 
seen in figure \ref{fig:xy}. We compare in figure \ref{fig:coll_polar_e} the radial collision profile to the radial structure seen in the e1m2, e2m2 and 
e3m2 simulations. In the lowest forced eccentricity case (e1m2), the pileups near the edge of the 2:1 MMR line up with the boundaries of the dip 
feature seen in the radial collision profile. The locations of the edges of the resonances are calculated using equations \ref{eq:res_fo} and 
\ref{eq:res_so}. The pileups not as pronounced near the other resonances. As discussed previously, the planetesimal density enhancements and gaps 
seen in polar coordinates follow the position of the giant planet in its orbit, which is shown by the vertical line. For the e2m2 case, the density 
enhancements take up a much larger radial width, which smears out the corresponding colli- sional dust profile. In the highest forced eccentricity case 
(e3m2), the density enhancements occur well within an annulus containing the resonant region, completely al- tering the radial peak-valley dust 
morphology.

@@

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We 
attribute this to the more comlex symmetry, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 
2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:2 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance 
space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would 
require much higher resolution observations. For these reasons, we will focus on the 2:1 MMR as the main diagnostic indicator.

@@

Because the strength\footnote{More specifically, the strength of the resonant term in the disturbing potential} of the 2:1 resonance scales with $e_{g}$ and  $m_{g}$, a sufficiently low mass and low eccentricity planet will not create 
enough of a perturbation in the planetesimal disk to produce a detectable bump or dip feature. The colored contours in figure \ref{fig:bump_dip_diag} 
indicate the relative strength of the resonance as these quantities are varied. Although the resonance strength is equally sensitive to changes in 
mass and eccentricity, the peri - apocenter overlap of the resonance edges is more sensitive to changes in eccentricity. This is because changes to 
eccentricity affect both the resonance width and the forced eccentricity. As we showed in figure \ref{fig:coll_polar_m}, changing $m_{g}$ by a factor 
of 4 produced very minimal changes to the resulting dust profile.

@@

Compared to using the corotationg gap opened by the planet as a diagnostic \citep{2013ApJ...777L..31D, 2016ApJ...820...29D}, measuring 
variations in the dust emission near MMRs is much more subtle and requires much higher spatial resolution and sensitivity. As we have shown, the 
bump vs dip feature near the 2:1 MMR is marginally detectable with ALMA for the nearest protoplanetary disks. Another complicating factor is that 
the inner $\sim$ 10 AU of most protoplanetary disks are optically thick in the sub-mm. The solution to this problem is to instead observe at longer 
wavelengths where the inner disk is optically thin. This comes at the expense of much poorer resolution. However, future radio facilities like the NG-
VLA are expected to achieve sub-AU resolution for nearby planet-forming disks \citep{2018ASPC..517..147R}. In the more near-term, techniques 
like Gaussian process fitting present a promising way to recover substructure at sub-beam resolution \citep{2020arXiv200507709J}.

@@

\section{Appendix: Nonaxisymmetric Structure in Boley 2017}\label{sec:boley_plot}

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/boley_rtheta}
    \caption{Positions of planetesimals at the end of the simulation presented in \citet{2017ApJ...850..103B} in polar coordinates. The solid line marks the location of a 50 $M_{\bigoplus}$ planet embedded in the disk. The dashed lines mark the locations of first order MMRs. A second planet is also present at 32.3 AU, although it is not massive enough to induce any significant structure in the disk near associated resonances.\label{fig:boley_rtheta}}
\end{figure*}

>>>>

\noindent where $\lambda = \varpi + M$ is the mean longitude of a
body, with $M$ being the mean anomaly. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citep{2000ssd..book.....M}. An important 
point to note, which we will come back to later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of 
the $\phi$. For second and higher order resonances, the maximum libration width is given by equation 8.58 in \citet{2000ssd..book.....M} as

@@

This also suggests that the prominent bumps and dips seen in figure \ref{fig:coll_hist_r} are mainly produced by bodies outside of resonance, although 
the resonant population does make some of the features more pronounced. To further understand this, we consider the nonaxisymmetric structure 
seen in figure \ref{fig:xy}. We compare in figure \ref{fig:coll_polar_e} the radial collision profile to the radial structure seen in the e1m2, e2m2 and 
e3m2 simulations. In the lowest forced eccentricity case (e1m2), the pileups near the edge of the 2:1 MMR line up with the boundaries of the dip 
feature seen in the radial collision profile. The locations of the edges of the resonances are calculated using equations \ref{eq:res_fo} and 
\ref{eq:res_so}. The pileups not as pronounced near the other resonances. As discussed previously, the planetesimal density enhancements and gaps 
seen in polar coordinates follow the position of the giant planet in its orbit, which is shown by the vertical line. For the e2m2 case, the density 
enhancements take up a much larger radial width, which smears out the corresponding collisional dust profile. In the highest forced eccentricity case 
(e3m2), the density enhancements occur well within an annulus containing the resonant region, completely al- tering the radial peak-valley dust 
morphology.

@@

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We 
attribute this to the more complex symmetry, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 
2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:2 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance 
space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would 
require much higher resolution observations. For these reasons, we will focus on the 2:1 MMR as the main diagnostic indicator.

@@

Because the strength\footnote{More specifically, the strength of the resonant term in the disturbing potential} of the 2:1 resonance scales with $e_{g}$ and  $m_{g}$, a sufficiently low mass and low eccentricity planet will not create 
enough of a perturbation in the planetesimal disk to produce a detectable bump or dip feature. The colored contours in figure \ref{fig:bump_dip_diag} 
indicate the relative strength of the resonance as these quantities are varied. Although the resonance strength is equally sensitive to changes in 
mass and eccentricity, the peri-apocenter overlap of the resonance edges is more sensitive to changes in eccentricity. This is because changes to 
eccentricity affect both the resonance width and the forced eccentricity. As we showed in figure \ref{fig:coll_polar_m}, changing $m_{g}$ by a factor 
of 4 produced very minimal changes to the resulting dust profile.

@@

Compared to using the corotating gap opened by the planet as a diagnostic \citep{2013ApJ...777L..31D, 2016ApJ...820...29D}, measuring 
variations in the dust emission near MMRs is much more subtle and requires much higher spatial resolution and sensitivity. As we have shown, the 
bump vs dip feature near the 2:1 MMR is marginally detectable with ALMA for the nearest protoplanetary disks. Another complicating factor is that 
the inner $\sim$ 10 AU of most protoplanetary disks are optically thick in the sub-mm. The solution to this problem is to instead observe at longer 
wavelengths where the inner disk is optically thin. This comes at the expense of much poorer resolution. However, future radio facilities like the NG-
VLA are expected to achieve sub-AU resolution for nearby planet-forming disks \citep{2018ASPC..517..147R}. In the more near-term, techniques 
like Gaussian process fitting present a promising way to recover substructure at sub-beam resolution \citep{2020arXiv200507709J}.

@@

\setcounter{secnumdepth}{0}
\section{Appendix: Nonaxisymmetric Structure in Boley 2017}\label{sec:boley_plot}

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{figures/boley_rtheta.png}
    \caption{Positions of planetesimals at the end of the simulation presented in \citet{2017ApJ...850..103B} in polar coordinates. The dashed line marks the location of a 50 $M_{\bigoplus}$ planet embedded in the disk. First order MMRs with the planet are indicated with arrows. A second planet is also present at 32.3 AU, although it is not massive enough to induce any significant structure in the disk near associated resonances.\label{fig:boley_rtheta}}
\end{figure*}
